---
title: "Untitled"
author: "Mauricio Mardones I."
date: "`r Sys.Date()`"
output: pdf_document
---


Install pkgs from [Iago Github](https://github.com/flr/ss3om) 

```{r setup1, echo=FALSE}
set.seed(999)
rm(list = ls())
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

```{r}
# instal pkgs to read SS3 out puts
#install.packages("ss3om", repos=structure(
#  c(CRAN="https://cloud.r-project.org/", FLR="https://flr-project.org/R")))
library(ss3om)
library(here)
library(FLasher)
# viz
library(ggthemes)
library(egg)
library(FLfse)
```


```{r}
dir <- here("test")
stk <- readFLSss3(dir)
plot(stk)
stk_stf <- window(stk, end = 2020) # cut off data after 2017
stk_stf <- stf(stk_stf, nyears = 10) # add 10 years
range(stk_stf)
```

Ploting variables from `SS3`krill stock assesment model

```{r}
plot(stk_stf) + 
  geom_vline(xintercept = 2020) +
  theme_few()
```
```{r}
### get stock
stk_par <- as(stk_stf, "FLStock")

plot(stk)
### year 1 is an unfished stock

### get stock-recruitment model
sr <- FLSR(params = params(stk_par),
           model = model(stk_par))
model(sr)
params(sr)
```
```{r}
sr <- as.FLSR(stk, model = "segreg") # create FLSR
sr <- fmle(sr) # fit model
plot(sr)


ctrl <- data.frame(year = 2023,
                   quant = "catch",
                   value = 5000)
ctrl <- fwdControl(ctrl)

stk_fwd <- fwd(stk_stf, 
               control = ctrl, 
               sr = sr)
plot(stk_fwd) +
  geom_vline(xintercept = 2022)



ctrl <- data.frame(year = 2020:2032,
                   quant = "f",
                   value = seq(from = 0.5, to = 0, length.out = 13))
ctrl <- fwdControl(ctrl)
stk_fwd <- fwd(stk_stf, control = ctrl, sr = sr)
plot(stk_fwd) + 
  geom_vline(xintercept = 2017.5)+
  theme_bw()
  

```

```{r}
ctrl <- data.frame(year = 2023,
                   quant = "catch",
                   value = 5000)
ctrl <- fwdControl(ctrl)

stk_fwd <- fwd(stk_stf, 
               control = ctrl, 
               sr = sr)
plot(stk_fwd) +
  geom_vline(xintercept = 2022)

ctrl <- data.frame(year = 2020:2032,
                   quant = "f",
                   value = seq(from = 0.5, to = 0, length.out = 13))
ctrl <- fwdControl(ctrl)
stk_fwd <- fwd(stk_stf, control = ctrl, sr = sr)
plot(stk_fwd) + 
  geom_vline(xintercept = 2017.5)+
  theme_bw()


f_trgts <- seq(from = 0, to = 1, by = 0.1)
stks <- lapply(f_trgts, function(x) {
  ctrl <- data.frame(year = 2020:2032,
                     quant = "f",
                     value = x)
  ctrl <- fwdControl(ctrl)
  stk_fwd <- fwd(stk_stf, control = ctrl, sr = sr)
  return(stk_fwd)
})
stks <- FLStocks(stks)
names(stks) <- f_trgts
plot(stks)+
  theme_bw()+
  scale_color_viridis_d()
```



```{r results = "hide", message = FALSE}
library(FLCore) # core functionality
library(FLasher) # for projections
library(ggplotFL) # for plotting
```

### Prepare operating model (OM)

`stk` is the basis for the operating model. Extend the stock for 10 years (without uncertainty):

```{r}
stk_fwd <- window(stk, end = 2022) 
stk_fwd <- stf(stk_fwd, nyears = 10) # add 10 years
```

And create a recruitment model (without any error)

```{r results = "hide", warning = FALSE}
sr <- as.FLSR(window(stk, start = 2004, end = 2021), 
              model = "segreg") # create FLSR
sr <- fmle(sr, control = list(trace = 0)) # fit model
# trace = 0 turns off console output of model fitting
```

Some recruitment and SSB values from the beginning/end of the time series are cut off in `sr` to account for the recruitment age (might be different from 0) and the time lag between SSB the recruitment the next year.

### Management procedure (MP) - chr rule

First, we use a constant harvest rate approach. A harvest rate is the catch divided by the stock size. We will use a harvest control rule (HCR) similar to ICES' chr rule, one of the approaches for category 3 data-limited stocks.

This HCR sets the catch $C$ by multiplying a stock size indicator $I$ with a target harvest rate $H$:

$$
C = H \times I \times b
$$

where $b$ is a biomass safeguard, that reduces the catch when the stock size $B$ falls below a trigger level $I_{\text{trigger}}$:

$$
b = \text{min}\left(1, I/I_\text{trigger} \right)
$$

We define some (arbitrary) control paramters for this HCR:
```{r}
Htrgt <- 0.2
Itrigger <- 50
```

#### MSE loop chr rule

Then we can set up an MSE loop. This follows the typical ICES assessment schedule, where the assessment is conducted in an intermediate year (`ay`), data is available up to the year before the intermediate year (`ay-1`), and catch advice is provided for the year following the intermediate year (`ay+1`). Consequently, we start the MSE loop in year 2017 (the intermediate year), to set the first catch advice for 2018.

```{r, warning = FALSE, message = FALSE}
### loop through years
for (ay in 2022:2031) {
  
  ### OM
    ### OM - observations: assume perfect observations
    ### data available up to year ay-1
    stk0 <-  window(stk_fwd, end = ay - 1) ### observed stock
  
  ### MP
    ### MP - estimator (stock assessment): none - assume perfect knowledge
    stk0 <- stk0 ### modelled stock
    I <- tail(ssb(stk0)) ### stock size indicator
  
    ### MP - decision model
    ### hockey-stick constant harvest rate (chr rule)
    
    ### biomass safeguard: SSB status relative to trigger
    b <- ifelse(I >= Itrigger, 1, I/Itrigger)
    
    ### calculate catch advice
    advice <- c(Htrgt * I * b)
  
  ### OM
    ### OM - Implementation model: assume perfect implementation
    catch <- advice
  
    ### OM - fishery & stock dynamics
    ### set catch target
    ctrl <- fwdControl(data.frame(year = ay + 1,
                                  quant = "catch",
                                  value = catch))
    ### projection (update OM by one year)
    stk_fwd <- fwd(stk_fwd, control = ctrl, sr = sr)
    
  ### status message
  cat(paste0("finished year ", ay, ", MP target C=", round(catch), "\n"))
    
}

```

Then, we can plot the operating model results.

```{r, dpi=300}
plot(stk_fwd)
```

Metrics such as mean fishing mortality can be seen with the usual functions.

```{r}
fbar(stk_fwd)
ssb(stk_fwd)
```

### Management procedure (MP) - ICES MSY rule

Second, we will use the ICES MSY rule. This HCR is the standard approach for ICES data-rich category 1 stocks and works very similar to the chr rule above, but targets a fishing mortality $F$:

$$
F = F_\text{target} \times \text{min}\left(1, B/B_\text{trigger} \right)
$$

We can define the control parameters of the MSY rule:

```{r}
Ftrgt <- 2.31
Btrigger <- 15
```

#### MSE loop ICESMSY rule

The output of the ICES MSY rule is a fishing mortality. However, this needs to be converted to catch value with a short-term forecast.

The ICES MSY rule is typically applied based on the results from a stock assessment, and a short-term forecast. In this example, we bypass the stock assessment and pass the stock size directly to the MP. It is therefore a shortcut MSE.

```{r, warning = FALSE, message = FALSE, dpi=300}
### loop through years
for (ay in 2022:2031) {
  
  ### OM
    ### OM - observations: assume perfect observations
    stk0 <-  window(stk_fwd, end = ay - 1) ### observed stock
  
  ### MP
    ### MP - estimator (stock assessment): none - assume perfect knowledge
    stk0 <- stk0 ### modelled stock
  
    ### MP - decision model
    ### hockey-stick advice rule
    ### SSB status relative to Btrigger (control parameter)
    mult <- ifelse(tail(ssb(stk0)) >= Btrigger, 1, tail(ssb(stk0))/Btrigger)
    ### set F target
    target <- Ftrgt * mult
  
    ### convert F target into catch advice with short-term forecast
    ### current estimated F
    fsq <- c(tail(fbar(stk0))) 
    ### recruitment assumption (in decision model): average of last 10 years
    ### format as FLPar object to tell FLR this is a parameter value
    rec_average <- FLPar(mean(tail(rec(stk0), 10)))
    ### create recruitment model
    sr0 <- FLSR(model = "geomean", 
                residuals = FLQuant(1), 
                params = rec_average)
    ### extend stock used in short-term forecast 
    stk0_stf <- stf(stk0, nyears = 2)
    ### set short-term forecast target
    ### (assume status quo F for intermediate year, then F target)
    ctrl0 <- fwdControl(data.frame(year  = c(ay, ay + 1),
                                   quant = c("f", "f"),
                                   value = c(c(fsq), c(target))))
    ### run short-term forecast
    stk0_stf <- fwd(stk0_stf, control = ctrl0, sr = sr0)
    ### extract catch advice
    advice <- c(tail(catch(stk0_stf)))
  
  ### OM
    ### OM - Implementation model: assume perfect implementation
    catch <- advice
  
    ### OM - fishery & stock dynamics
    ### set catch target
    ctrl <- fwdControl(data.frame(year = ay + 1,
                                  quant = "catch",
                                  value = catch))
    ### projection (update OM by one year)
    stk_fwd <- fwd(stk_fwd, control = ctrl, sr = sr)
    
  ### status message
  cat(paste0("finished year ", ay, ", MP target F=", 
             round(ctrl0$value[2], 2), "\n"))

}
  
```


```{r, warning = FALSE, message = FALSE, dpi=300}
plot(stk_fwd)+
  geom_line()+
  theme_few()
```
###


```{r}
### fit SAM to cod
fit <- FLR_SAM(stk = stk_fwd, 
               idx = cod4_idx, 
               conf = cod4_conf_sam)
### the output is a normal SAM object
fit
tail(summary(fit))
plot(stk)
```



```{r}
idx <- readRDS(here("FLR_examples",
            "7_FLR_MSE","OM_files",
            "idx.rds")) ### indices
catch_res <- readRDS(here("FLR_examples",
            "7_FLR_MSE","OM_files",
            "catch_res.rds")) ### catch residuals
idx_dev <- readRDS(here("FLR_examples",
            "7_FLR_MSE","OM_files",
            "idx_dev.rds")) ### catch residuals
### load projection method
source(here("FLR_examples",
            "7_FLR_MSE",
            "MSE_functions",
            "funs_MSE_oem.R"))
### create the FLoem
oem <- FLoem(observations = list(stk = stk0, idx = idx),
             deviances = list(stk = FLQuants(catch.dev = catch_res),
                              idx = idx_dev),
             method = oem_WKNSMSE,
             ### additional arguments, passed to method
             args = list(idx_timing = c(0, -1), ### index timing rel. to ay
                         catch_timing = -1, ### catch timing rel to ay
                         use_catch_residuals = TRUE, ### add catch noise? 
                         use_idx_residuals = TRUE ### add idx noise?
                         ))

class(oem)
slotNames(oem)
```

