---
title: "Black Spott Seabream: Short-Cut MSE approach for robustness tests of harvest control rules in sex-structured models"

author: 
 - Henning Winker (GFCM) 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    #keep_tex: false
vignette: >
  %\VignetteIndexEntry{ss3daigs Vignetted}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r wrap-hook, echo = FALSE,warning=FALSE,message=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```


```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, 
                      comment = "  " ,
                      fig.align = 'center',
                      cache=FALSE,
                      tidy.opts=list(width.cutoff=55), 
                      tidy=TRUE)
```

Load packages


```{r, message=FALSE,warning=FALSE}
# Load
library(ggplot2)
library(FLCore)
library(ggplotFL)
library(mse)
library(FLRef)
library(ggpubr)
library(mseviz)
library(r4ss)
library(doParallel)
library(here)
```

# Short-Cut MSE for Harvest Control Rule (HCR) 

In contrast to a full Management Strategy Evaluation (MSE) simulation design (Punt et al. 2017), the MSE  ‘shortcut’ approach, omits the step of the annual updating of the estimation model (assessment) in the feedback control.  Instead, it emulates an annual update of the benchmark assessment model by passing outcomes ($SSB$ and $F$) from the 'true' age-structured dynamics from the operating model (OM) with assessment error to the harvest control rule (HCR) and catch implementation system. 

The HCRs were implemented using a simulated feedback control loop between the implementation system and the operating model, where the implementation system translates the emulated assessment outcome via the HRC into  the Total Allowable Catch (TAC) advice. The feedback control loop between the implementation system and the OM allows accounting for the lag between the last of year data used in the assessment and the implementation year of catch advice ($C_{adv}$). 

For blackspot seabream, the implementation system of the harvest control rule is based on the assumption that advice is given for year $y+1$ based on an assessment completed in year $y$, which is fitted to data up until last data year $y-1$. Therefore implementation of the  derived $C_{adv}$ through HCR requires projection of the stock dynamics by way of a short-term forecast. To do this, numbers-at-age were projected through the year of assessment. Status quo recruitment, $M_a$, $w_a$ and $mat_a$ were set as the mean of the last 3 years. A projection based on a fixed fishing mortality-at-age to the last year ($y-1$) in the assessment is then made through to the implementation year ($y+1$).

The limitations of the MSE short-cut approach are that it cannot fully account for uncertainties resulting from imperfect sampling of the full age-structure (e.g. poorly sampled recruits), observation error, misspecified model assumptions and selectivity. On the other hand, the short-cut MSE approach is straight-forward to implement (`FLR`) and reduced complexity and computation time when the focus is predominantly optimizing HCRs for setting quotas on the premises that a benchmark assessment form the basis for the advice.

Here, the MSE short-cut approach is implemented using the tools available in the Fisheries Library for R (FLR; Kell et al., 2007; https://flr-project.org/) 


## Glossary

The following glossary summarizes key HCR parameters and associated target and limit reference points
that are considered for tuning the candidate HCRs to optimise the trade-offs between maximising fishing
opportunity and risk:


+ $F_{MSY}$: target reference point for fishing mortality at Fmsy (or its proxy), (e.g. $F_{B35}$)

+ $B_{MSY}$: the average biomass around which the biomass fluctuated when fishing at $F_{MSY}$ or its proxy (e.g. $B_{35}$)

+ $B_{lim}$: a deterministic biomass limit reference point below which a stock is considered to have reduced reproductive capacity. Here $B_{lim}$ was set to $0.25B_{tgt}$ 

+ $B_{pa}$: a precautionary biomass reference point set with high probability that biomass is above $B_{lim}$, which acts as a safety margin below which the risk of reduced reproductive capacity is increasing. When the biomass is estimated to be above Bpa, the stock is considered to be within safe biological limits in terms of its reproductive capacity. 

+ $C_{adv}$: advised catch as output of the management procedure 

+ $B_{trigger}$: biomass trigger point of the HCR, specified as change point of biomass below which fishing mortality reduced relative to Ftgt. Btrigger is typically specified as ratio to $B_{MSY}$.


# Build FLStock

SS3 outputs are loaded with the `readFLSss3()` into an `FLStock` object. The folder that contains the model outputs has to be specified.

In the following, the area outside is evaluated first.



```{r, eval=FALSE}
dir1 <- here("s1.4")
run = "krill.stock.wAP"
stk = window(ss3om::readFLSss3(dir1))
stk = simplify(stk)
# Fill NAs
stk@m.spwn[] = 0
stk@harvest.spwn[] = 0 
sr = ss3om::readFLSRss3(dir1,run)
stk@name = run
stk@desc = "2024, CCAMLR, SS3"
out = ss3om::readOutputss3(dir1)
```

```{r, echo=F}
run = "krill.stock.wAP"
load(file=file.path("rdata",paste0(run,".rdata")),verbose = T)
stka = stk
```

## Retune Stock-Recruitment


```{r}
# Extract SR pars
s = params(sr)[[1]]
R0 = params(sr)[[2]]
B0 = params(sr)[[3]]

# single sex
sr1 = srrTMB(as.FLSR(stka,
                     model=bevholtSV),
             spr0=mean(spr0y(stk)),
             r0.pr=c(R0,0.0001),
             s=s,
             s.est=F)
```


```{r , message=FALSE,warning=FALSE,fig.width=9,fig.height=7, fig.cap = paste0("Comparison of stock stock-recruitment function from ss3 and retuned single-sex with FLSRTMB ") }
plotsrs(FLSRs(ss3=sr,
              sr=sr1))
```
Check that the fbar range is consistent with `ss.starter` input.

```{r}
range(stk)
```

```{r , message=FALSE,warning=FALSE,fig.height=6,fig.width=8, fig.cap = paste0("Seasonal  stock trajectories") }
plot(stk,
     metrics=list(SSB=function(x)unitSums(ssb(x)[,,,1]),
                  F=function(x)fbar(x),
                  Catch=function(x)catch(x),
                  Rec=function(x)unitSums(rec(x))))+
  theme_bw()+
  ylab("F")+
  xlab("Year")+
  facet_wrap(~qname,
             scales="free_y")
```

\pagebreak

## Plot SS3 Stock Dynamics


```{r fig1.1.2, fig.height=7,fig.width=7, fig.cap = paste0("Stock assessment trajectories at age")}

plotdyn(stk)+
  theme_bw()+
  scale_color_viridis_d(option="G")
```

\newpage 

```{r fig1.1.3, fig.height=6, fig.cap = paste0("Stock biology trajectories at age")}

plotbioyr(stk)+
  ggtitle(paste0(stk@name))+
  scale_color_viridis_d(option="G")

```

\newpage 

```{r fig1.1.4, fig.height=6, fig.cap = paste0("Annual stock quanties at age")}
plotbioage(stk)+
  theme(legend.position = "none")+
  scale_color_viridis_d(option="G")
```

\pagebreak

## Consistency checks using backtesting


```{r}
set.seed(123)
```

Get bias adjusted recruitment deviations from ss3 model

Simplify to annual sex-structured model

```{r}
if(dims(stk)$season>1){
stka = simplify(stk,'season',weighted = TRUE,harvest=TRUE)

discards.wt(stka) = stock.wt(stka)
stka@discards  =  computeDiscards(stka)
# Make annual sra
sra = sr
params(sra) = FLPar(an(sr@params[,1]),params=rownames(sr@params))
} else {
sra = sr1
stka = stk
}
```


```{r}
yrs = an(dimnames(stk)$year)
recruit = out$recruit[out$recruit$Yr%in%yrs,]
dms <- list(year = yrs)
sigR =  mean(an(out$sigma_R_info[1:2,"SD_of_devs_over_sigma_R"])) # Realised sigR
residuals <- FLQuant(exp(recruit$dev - 0.5 * recruit$biasadjuster *sigR^2), 
        dimnames = c(age = 0, dms), units = "")
recs = FLQuant(recruit$pred_recr, dimnames = c(age = 0, dms), units = "")


if (dims(stk)$unit == 2) recs <- expand(recs, unit = c("F", "M"))

```

```{r}
if (dims(stka)$unit == 3)
yrs = an(dimnames(stka)$year)
testC = fwd(stka,sr=recs[,ac(yrs[-1])],
  control=fwdControl(year=yrs[-1], value=(unitSums(catch(stka)[, ac(yrs[-1])])),
  quant="catch"))

testF = fwd(stka, sr=recs[,ac(yrs[-1])],
  control=fwdControl(year=yrs[-1], value=unitMeans(fbar(stka)[, ac(yrs[-1])]),
  quant="fbar"))
```

```{r hc1, message=FALSE,warning=FALSE,fig.width=9,fig.height=7, fig.cap = paste0("Comparison of stock trajectories from ss3om and a backtest under the same Fbar") }

plot(window(FLStocks(ss3om=stka,
                     backtestC=testC,
                     backtestF=testF)))+
  theme_bw()+
  facet_wrap(~qname,scale="free_y")
```

Note that minor deviations are likely due to difficulties in precisely adjusting the rec devs with bias correction.

## Estimate candidate reference points

```{r}

# Main recdevs
recyrs = recruit$Yr[recruit$era =="Main"]
maindevs = unitSums(residuals[,ac(recyrs)])
rho = cor(maindevs [,-1],maindevs [,-length(maindevs)])
sigmaR = out$sigma_R_in

rho 
sigmaR

# MSY refpts
Bmsy <- out$derived_quants$Value[out$derived_quants$Label=="SSB_MSY"]
Fmsy <- 0.001
MSY <- out$derived_quants$Value[out$derived_quants$Label=="Dead_Catch_MSY"]
out$derived_quants$Value[out$derived_quants$Label=="B_MSY/SSB_unfished"]
# Short cut devs
ay = out$endyr # assessment year
SSBcv <- out$derived_quants$StdDev[out$derived_quants$Label==paste0("SSB_",ay)]/
  out$derived_quants$Value[out$derived_quants$Label==paste0("SSB_",ay)]

Fcv <- out$derived_quants$StdDev[out$derived_quants$Label==paste0("F_",ay)]/
  out$derived_quants$Value[out$derived_quants$Label==paste0("F_",ay)]

```


# Tuning grid

Specify EQSIM outputs

```{r}
Blim=145000
Bpa=620000
Btri.eq = 620000
Feq =0.018  
Fp05.eq =0.018
```

Note that in this case the "true" $F_{MSY}$ as the property of the model is smaller than the $F_{MSY}$ derived from EQsim. This may be explained by the presence of the harvest control rule resulting in effective taking place, on average, below Btrigger.

```{r}
Fmsy # SS3
Feq
```

Function to find $B$ for $F$ at equilibrium

```{r}
fwdB4F = function(stock,sr,Fs=0.2,nfy=100){
    if (class(stock) == "FLStockR") {
      stock = as(stock, "FLStock")
    }
    fyrs = (dims(stock)$maxyear + 1):(dims(stock)$maxyear + nfy)
    nfy = length(fyrs)
    stkf = stf(stock, nfy)
    bx = do.call(c, lapply(an(Fs),function(x){
    ictrl = fwdControl(data.frame(year = fyrs, quant =  "fbar", value = x))
    out = fwd(stkf, sr = sr, control = ictrl)
    an(tail(unitSums(ssb(out))))
    }))
    
    data.frame(F=an(Fs),B=bx)
}

```



```{r}
Fx = c(rev(seq(0.5,1,0.025)))
Ftgt = FLPar(c(Fmsy,Fx*Feq),
             params=c("Fmsy.om",
                      paste0(Fx,".Feq"))) 
Ftgt
```

```{r,eval=F}
Bftgt = fwdB4F(stka,
               sra,
               Fs=Ftgt,
               nfy=200)
Bftgt
```

```{r, eval=FALSE, echo=FALSE}
save(Fx,
     Ftgt,
     Bftgt,
     file="rdata/ftune.sbr.rdata")
```

```{r,echo=F}
load(file="rdata/ftune.sbr.rdata")
```

```{r}
Ftgt.tune = FLPar(c(Bftgt$F,Bftgt$B),
                  params=c("Fmsy.om",
                           paste0("Feq",Fx),
                           "Bmsy.om",
                           paste0("Bftgt",Fx))) 
```



```{r}
np = an(nrow(Ftgt.tune))

df = data.frame(
Tune = rownames(Ftgt.tune)[1:(np/2)],
Ftgt = round(an(Ftgt.tune)[1:(np/2)],3),
Btrigger = round(Btri.eq,1),
Btgt = round(an(Ftgt.tune)[(np/2+1):np],1),
"xB0" =  round(an(Ftgt.tune)[(np/2+1):np]/B0,3))
df$Btrigger[1] = 0
df$xB0 = round(df$xB0,2)
```

```{r}
knitr::kable(df,"pipe",
      align ="lccccc",
      caption="Option: Initial tuning grid with EQSIM Btrigger based on Feq tuning.")

```


```{r}
refpts = FLPar(Fmsy=Fmsy,
               Feq=Feq,
               Bmsy=Ftgt.tune["Bmsy.om"],
               Beq=Ftgt.tune["Bftgt1"],
               Blim = Blim,
               Bpa=Bpa,
               Btrigger=Btri.eq,
               B0=B0,
               R0=R0)

refpts
```
Create `FLStockR` with `@refpts`

```{r}
stkr = FLStockR(stka)
stkr@refpts = refpts
```

Summarize short-cut params

```{r}
spars = FLPar(s=s,
              sigmaR=sigmaR, 
              rho=rho,
              Fcv=Fcv,
              SSBcv=SSBcv)
```

```{r , message=FALSE,warning=FALSE,fig.height=7,fig.width=9, fig.cap = paste0("Status Advice plot  showing stock trajectories of Recruitment, $SSB$, $F$, recruitment and $Yield$") }

plotAdvice(stkr)
```

```{r eval=FALSE}
save(stk,
     stka,
     stkr,
     refpts,
     sr,
     sra,
     spars,
     file="rdata/om.ss3ref.sbr.rdata") 
```

# Set up short-cut MSE

Load OM conditioned to Stock Synthesis

```{r}
load("rdata/om.ss3ref.sbr.rdata",verbose=T)
```


Next set up the MSE horizon

```{r}
# data year
dy <- dims(stka)$maxyear
# FINAL year
fy <- dy+10
# assessment year
ay = dy+1
# intermediate years
iy = ay

```

For illustration the number of iterations are reduced to 100.

```{r}
# NUMBER iterations
it <- 100
```

Subset the 1000 simulated stock iterations to the first 100

```{r}
stki = propagate(stka,
                 it)
```

Generate recruitment 

```{r}
srdevs <- rlnormar1(n=it,
                    sdlog=spars["sigmaR"],
                    rho=spars["rho"], 
                    years=seq(dy, fy))
# Sex-structured
if (dims(stki)$unit == 2) srdevs  <- expand(srdevs ,
                                            unit = c("F", "M"))

```


Now construct the `FLom` object from the `mse` package by passing on `FLStock`,`refpts`, `sr` and the method used for foward projections.

```{r}
om <- FLom(stock=stki,
           refpts=refpts, 
           sr=sra,
           projection=mseCtrl(method=fwd.om),
           deviances=srdevs)

om
```

Next add the structure for the future years: average of last 3 years

```{r}
om <- fwdWindow(om, 
                end=fy)
```

Next, a so called observation error is constructed. In the case of the short-cut MSE, it simply holds the "perfect" stock information. For a full MSE with inbuilt estimation model  it would also generate the observations with errors, such a catch-at-age and survey numbers at age for SAM or a4a, or biomass surveys indices and catches for SPicT or JABBA.  

```{r tidy=FALSE}
oem <- FLoem(
  observations=list(stk=stock(om)),
  method=perfect.oem
)
```

However, there is increasing realisation that the assessment estimates are imperfect. Therefore, ICES has implemented procedures to add uncertainty about the key quantities $F$ and $SSB$, where the error on $F$ is specified by a the random error term $Fcv$ and a first order autocorrelation parameter $Fphi$ and the precision of $SSB$ can specified by $SSBcv$ 


Short-cut deviations

```{r tidy=FALSE}
sdevs <- shortcut_devs(om, 
                       Fcv=spars["Fcv"], 
                       Fphi=0.5, 
                       SSBcv=spars["SSBcv"])

```

Finally, the implementation error module `iem` is setup. In this case, with a random catch implementation error of 10\%.

```{r}
iem <- FLiem(method=noise.iem,
  args=list(noise=rlnorm(it, rec(om) %=% 0, 0.1)))
```


```{r,eval=FALSE}
save(om, 
     oem,
     sdevs,
     iem,
     file="rdata/flom.sbr.rda")
```

```{r,echo=FALSE}
load(file="rdata/flom.sbr.rda")
```

## Setting up harvest control rules



This can be effectively implemented ICES advice rule hockey-stick by setting the $B_{trigger}$ to zero, using the `icesControl` function. This function can also take the `SSBdevs` and `Fdevs` that implement the deviations from the SSB and F with aim to account for assessment errors.


Adjusted mse controls for sex-structured stocks

```{r}
shortcut.sa2 <- function(stk, 
                         idx, 
                         SSBdevs=unitSums(ssb(stk)) %=% 1, 
                         args, 
                         tracking, 
                         ...) {
  # DIMS
  y0 <- args$y0
  dy <- args$dy
  ay <- args$ay
  it <- args$it

  # SUBSET oem stock
  stk <- window(stk, end=dy)

  ind <- FLQuants(
    # SSB + devs 
    #><> add unitSums
    ssb=unitSums(ssb(stk)) * window(SSBdevs, 
                                    start=y0, 
                                    end=dy))

  track(tracking, "conv.est", ac(ay)) <- 1

  list(stk=stk, ind=ind, tracking=tracking)
}



tac.is2 <- function(stk, 
                    ctrl,
                    args, 
                    output="catch",
                    recyrs=-2,
  Fdevs=unitMeans(fbar(fut)) %=% 1, 
  dtaclow=NA, 
  dtacupp=NA, 
  fmin=0, 
  reuse=TRUE,
  initac=metrics(stk, output)[, ac(iy - 1)], tracking) { 
  # EXTRACT args
  spread(args)


  # SET control years
  cys <- seq(ay + management_lag, ay + management_lag + frq - 1)

  # PREPARE stk for cys, biology as in last nsqy years
  fut <- fwdWindow(stk, end=cys[length(cys)], nsq=nsqy)

  # PARSE recyrs if numeric
  id <- dimnames(stk)$year

  # COERCE to list
  if(!is.list(recyrs)) {
    recyrs <- list(recyrs)
  }
  
  # PARSE list
  for(i in recyrs) {
    if(is(i, 'character')) {
      id <- id[!id %in% i]
    } else if(all(i < 0)) {
      if(length(i) == 1)
        id <- rev(rev(id)[-seq(abs(i))])
      else
        id <- rev(rev(id)[i])
    } else if(all(i > 0)) {
      id <- rev(rev(id)[seq(abs(i))])
    }
  }

  # SET years to use
  recyrs <- id

  # CHECK recyrs
  if(!all(recyrs %in% dimnames(stk)$year)) {
    stop("'recyrs' cannot be found in input stk")
  }

  # TODO: OTHER rec options
  
 
  # SET GM recruitment from past
  #><> add unitSums()
  gmnrec <- exp(yearMeans(log(unitSums(rec(stk))[, recyrs])))

  # SETUP SRR
  srr <- predictModel(model=rec~a, params=FLPar(a=gmnrec))

  # STORE geomeanrec value 
  track(tracking, "gmrec.isys", ay + management_lag) <- gmnrec

  # ADD F deviances for 1 year

  # reuse = TRUE
  if(isTRUE(reuse) | toupper(reuse) == 'F') {
    ftar <- rep(c(ctrl[1,]$value * Fdevs[, ac(cys[1])]), length(cys))
  # reuse = FALSE
  } else {
    ftar <- c(ctrl$value * Fdevs[, ac(cys)])
  }

  # TRACK Ftarget
  track(tracking, "fbar.isys", cys) <- ftar

  # FORECAST for iyrs and my IF mlag > 0,
  if(management_lag > 0) {
 
    # SET F for intermediate year #><> added unitMeans
    #><> add unitMeans()
  
    fsq <- unitMeans(fbar(stk))[, ac(dy)]

    # TODO: ADD TAC option

    # CONSTRUCT fwd control
    fctrl <- fwdControl(
      # ay as intermediate with Fsq TODO: Other options
      list(year=seq(ay - data_lag + 1, length=management_lag),
        quant="fbar", value=rep(c(fsq), management_lag)),
      # target
      list(year=cys, quant="fbar", value=c(ftar))
    )

  # else only for my
  } else {
    fctrl <- fwdControl(
      list(year=ay + management_lag, quant="fbar", value=ftar))
  }

  # RUN STF ffwd
  fut <- ffwd(fut, sr=srr, control=fctrl)

  # ID iters where hcr set met trigger and F > fmin
  id <- c(tracking[[1]]["decision.hcr", ac(ay)] > 2) &
    c(unitMeans(fbar(fut))[, ac(ay + management_lag)] > fmin)

  # EXTRACT catches
  if(isTRUE(reuse) | toupper(reuse) == "C") {
    TAC <- expand(unitSums(catch(fut))[, ac(cys)[1]], year=seq(length(cys)))
  } else {
    TAC <- unitSums(catch(fut))[, ac(cys)]
  }

  # GET TAC dy / ay - 1
  if(ay == iy)
    prev_tac <- rep(c(initac), length=args$it)
  else
    prev_tac <- c(tracking[[1]]["isys", ac(ay)])

  # APPLY upper and lower TAC limit, if not NA and only for id iters
  if(!is.na(dtacupp)) {
    iter(TAC, id) <- pmin(c(iter(TAC, id)), prev_tac[id] * dtacupp)
  }
  if(!is.na(dtaclow)) {
    iter(TAC, id) <- pmax(c(iter(TAC, id)), prev_tac[id] * dtaclow)
  }

  # CONSTRUCT fwdControl
  # TODO: USE frq here
  ctrl <- fwdControl(lapply(seq(length(cys)), function(x)
    list(year=cys[x], quant=output, value=TAC[,x])))
    
  return(list(ctrl=ctrl, tracking=tracking))
}


```



```{r}
arule <- mpCtrl(list(
  # (est)imation method: shortcut.sa + SSB deviances
  est = mseCtrl(method=shortcut.sa,
    args=list(SSBdevs=sdevs$SSB)),

  # hcr: hockeystick (fbar ~ ssb | lim, trigger, target, min)
  hcr = mseCtrl(method=hockeystick.hcr,
    args=list(lim=0, trigger=0.01, target=Fmsy,
    min=0, metric="ssb", output="fbar")),

  # (i)mplementation (sys)tem: tac.is (C ~ F) + F deviance
  isys = mseCtrl(method=tac.is,
    args=list(recyrs=-2, fmin=0, Fdevs=sdevs$F))
  ))


```

This rule can now be run passing on the `om`, `oem` and `arule` and an additional argument to set the implementation year to 2024.

Note that the default setting assumes 1 year lag between data (reference) year and assessment (reporting) year and that the TAC is implemented the next year. In case the assessment is conducted in 2023, based on data from 2022 and the TAC is implemented for 2024.




```{r, eval=FALSE, echo=FALSE}
mseargs <- list(iy=dy,
                fy=fy, 
                data_lag=1, 
                management_lag=1, 
                frq=1)

load(file="rdata/mps.ftune.sbr1000.rdata",
     verbose=T)

mp.fixexFmsy = stks$Fmsy.OM

```


```{r,warning=FALSE}
mseargs <- list(iy=dy,
                fy=fy, 
                data_lag=1, 
                management_lag=1, 
                frq=1)

system.time(
run <- mp(om, 
          oem=oem, 
          ctrl=arule, 
          args=mseargs,
          verbose=T)
)

mp.fixexFmsy = run@om@stock

```


```{r fig9, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Initial OM and the MSE forecast horizon under a fixed $F_{tgt}$ rule")}

# make FLStocks from om until 2024 (implementation) and the run 
stk.fmsy = FLStocks(stock=window(om@stock,end=dy),
              fixedFmsy=mp.fixexFmsy)

plot(stk.fmsy)+
  facet_wrap(~qname,
             scales="free")+
  theme_bw()+
   geom_vline(xintercept = c(dy),
              linetype=2,col=1)+
  geom_vline(xintercept = c(fy),
             linetype=2,col=4)

```

The `run` can also be appended to the `om` to make a single `FLStockR` object with reference points

```{r}
runR = FLStockR(append(window(om@stock,
                              end=dy),
                       mp.fixexFmsy))
runR@refpts=refpts
```

This allows to quickly evaluate the stock status under a fixed $F_{MSY}$ rule.

It can be seen that despite "perfect" knowledge of the "true" $F_{MSY}$, and fishing pressure is on average $F_{MSY}$, the stock fails to attain biomass levels at $B_{MSY}$ with a relative high risk to fall below $B_{lim}$. This is a well known fact as a result of the lags between data and management and asymetric risks in that exceeding $F_{MSY}$ is more consequential on both $SSB$ and long term yield, then fishing below $F_{MSY}$, In the case of the latter, more biomass is left in the water, which provides increased future reproduction potential and catch opportunity.   

```{r, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Stock Status under a fixed $F_{tgt}$ short-cut MSE rule")}
thin = seq(1,it,5)
plotAdvice(iter(runR,thin))+
     geom_vline(xintercept = c(dy),
                linetype=2,col=1)+
    geom_vline(xintercept = c(fy),
               linetype=2,col=4)+
  scale_x_continuous(breaks=seq(1970,3000,5))+
  theme_bw()+
  theme(axis.text.x = element_text(size=8,
                                   angle=90,
                                   vjust=0.5))
```

Even relatively simplified short-cut MSE frameworks provide a powerful to explore alternative HCRs to achieve better trade-off between risks and yield. 

Here, the convential hockey-stick control rule is explored with different ratios of $F_{adv}$/$F_{tgt}$ and $B_{trigger}$/$B_{tgt}$ settings, where the $B_{trigger}$ promts a linear reduction in $F_{adv}$ if $SSB$ falls below it.


```{r , fig.height=5,fig.width=8,warning=FALSE,message=FALSE, fig.cap = paste0("Advice Rule from EQSIM")}

 plotGFCM(fadv = Feq,
          btrigger =Btri.eq,
          bthr = -1,
          ftgt = Feq,
          btgt=df$Btgt[2],
          blim=an(refpts["Blim"]),
          fmin=0,
          bclose=0,
          kobe  =FALSE,
          text=F,
          rel=F,
          xmax=2.5)
```


```{r , fig.height=12,fig.width=8,warning=FALSE,message=FALSE, fig.cap = paste0("Alternative HCR formulations, with $F_{min}=0$ (top) and $F_{min}=0.5F_{tgt}$ (bottom) under different $F_{adv}$ options set at $F_{adv}=F_{adv}$, $F_{adv}=0.8F_{adv}$ and at $F_{adv}=0.7F_{adv}$ from left to right")}

Fadv = c(df$Ftgt) 
Btgt = c(df$Btgt) 
Btri = c(df$Btrigger)

hcrs= Map(function(x,y,z){
p  = plotGFCM(fadv = x,
              btrigger =y,
              ftgt = x,
              btgt=z,
              blim=an(refpts["Blim"]),
              fmin=0,
              bclose=0,
              kobe  =FALSE,
              text=F,
              rel=F)
p = p
return(p)
},
x=Fadv,
y=Btri,
z=Btgt)
# plot ggplot list
ggarrange(plotlist=hcrs,
          nrow=6,
          ncol=4)

```

The same settings can be specified for the new `mps` function in `mse`, which allow to explore variations of the HCR parameters.

The function `combinations` enables to vary more than one parameter at the time. 

```{r}
hcrs =combinations(target=Fadv,
                   trigger=0)
hcrs$trigger = Btri # Overwrite
hcrs

save(om,
     oem,
     arule,
     hcrs,
     df,
     mseargs,
     file="rdata/inpMP.sbr.1000.rdata")
```



These changes in parameters can simply be passed on to the existing `arule` to run the variations with `mps`.  

```{r,eval=FALSE,warning=FALSE}
runs <- mps(om, 
            oem=oem,
            ctrl=arule, 
            args=mseargs,
            hcr=hcrs)
```

Note the this was run on a linux cluster in parallel

```{r}
# Divide en lotes de 100
ni <- length(hcrs$target)  # ← esto es clave

batch_size <- 100  # puedes ajustar este valor según tu iMac
batches <- split(seq(ni), ceiling(seq_along(seq(ni)) / batch_size))


for (b in seq_along(batches)) {
  runs_batch <- foreach(i = batches[[b]]) %dopar% {
    # mismo cuerpo del loop...
  }
  saveRDS(runs_batch, file = paste0("s1.4/runs_batch_", b, ".rds"))
}


runs <- list()
for (i in seq(ni)) {
  hcr = arule
  hcr$hcr@args$target = hcrs$target[i]
  hcr$hcr@args$trigger = hcrs$trigger[i]
  
  runi <- mp(om, oem=oem, ctrl=hcr, args=mseargs, verbose=TRUE, parallel = FALSE)
  runs[[i]] <- runi
}

```


```{r,eval=F}
library(doParallel)
load("rdata/inpMP.sbr.1000.rdata",
     verbose=T)
length(hcrs$target)
# Run in batches

ni = length(hcrs$target)
cl = ni

registerDoParallel(cl)
start = Sys.time()

runs <- foreach(i = seq(ni)) %dopar% {
  
  # set stock index
  hcr = arule
  hcr$hcr@args$target = hcrs$target[i]
  hcr$hcr@args$trigger = hcrs$trigger[i]
  
  runi <- mp(om, 
             oem=oem, 
             ctrl=hcr, 
             args=mseargs,
             verbose=T,
             parallel = FALSE)
  
  return(runi)
} # end of loop 
end = Sys.time()
time = end-start
time



nfs = length(df$Tune[-1])
tri = c(rep(".Btri.eq",
            each=nfs))
scenarios = c("Fmsy.OM",
              paste0(rep(c(df$Tune[-1]),1),tri))

names(runs) = scenarios

stks = FLStocks(lapply(runs,function(x){
  out = x@om@stock
  out = FLStockR(out)
  out@refpts = om@refpts
  out
}))

stkm = FLStocks(lapply(stks,function(x){
  stockMedians(x)
}))

plotAdvice(stkm)

save(stks,
     file=paste0("rdata/mps.ftune.sbr1000.rdata"))



```


Now combine with the Fixed $F_{MSY}$ run and see if we can do better.

```{r,eval=F}
nfs = length(df$Tune[-1])
tri = c(rep(".Btri.eq",each=nfs))
scenarios = c("Fmsy.OM",paste0(rep(c(df$Tune[-1]),1),tri))

names(runs) = scenarios

```

Save 



An easy way of basic plotting is to extract the `FLStocks` from the list of MSE runs with `lapply`

```{r,eval=F}
stks = FLStocks(lapply(runs,function(x){
  out = (x@om@stock)
  out = FLStockR(out)
  out@refpts = refpts[c(1:5)]
  out
  }))


```



```{r, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Initial OM and the MSE forecast horizon under a fixed $F_{tgt}$ rule and 6 HCR hockey-stick scenarios")}

ref = FLStockR(window(stock(om),end=dy))
ref@refpts = refpts
pstks = FLStocks(c(FLStocks(stock=window(ref,end=dy)),
              stks))

plotAdvice(iter(pstks,thin))+facet_wrap(~qname,scales="free")+
  theme_bw()+
  scale_color_manual(values=c("black",
                              ss3col(length(pstks))[-1]))+
  scale_fill_manual(values=c("darkgrey",
                             ss3col(length(pstks))[-1]))+
  geom_vline(xintercept = c(dy),
             linetype=2,col=1)+
    geom_vline(xintercept = c(fy),
               linetype=2,
               col=4)+
  scale_x_continuous(breaks=seq(1900,3000,5))+
  theme(axis.text.x = element_text(size=8, 
                                   angle=90,
                                   vjust=0.5))

```


```{r,eval=FALSE}
save(pstks,
     mseargs,
     file="rdata/shortcut_stks.sbr.rda",
     compress="xz")
```


```{r, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Median MSE forecast horizon under a fixed $F_{tgt}$ rule and 6 HCR scenarios")}

medstks = FLStocks(lapply(stks,function(x){
  stockMedians(x)
}))
medref = stockMedians(ref)
pmstks = FLStocks(c(FLStocks(stock=window(medref,end=dy)),
              medstks))

plotAdvice(pmstks)+facet_wrap(~qname,scales="free")+
  scale_color_manual(values=c("black",ss3col(length(pmstks))[-1]))+
  theme_bw()+xlab("Year")+
  geom_vline(xintercept = c(dy),linetype=2,col=1)+
    geom_vline(xintercept = c(fy),linetype=2,col=4)+
  scale_x_continuous(breaks=seq(1900,3000,5))+
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))

```



# Performance Evaluation with adjustment for sex-tructured stocks

## Define Metrics for performance evaluation of sex-structured 

```{r}
metrics <- list(SB = function(x)unitSums(ssb(x)), 
                F = function(x)unitMeans(fbar(x)), 
                C = function(x)unitSums(catch(x)),
                Rec= function(x)unitSums(rec(x)))
```


Define performance statistics - ordered alphabetically

```{r}
stats <- list(
  a.medianFmsy= list(~yearMedians(F/Fmsy), name="F/Fmsy",
    desc="Median annual F/Fmsy"),
  b.medianBmsy = list(~yearMedians(SB/Bmsy), name="B/Bmsy",
    desc="Median annual B/Bmsy"),
  c.medianCmsy = list(~yearMedians(C/MSY), name="Catch/MSY",
    desc="Median Catch/MSY over years"),
  d.aavC = list(~yearMedians(iav(C)), name="AAV",
    desc="Median annual variation in catches"),
  e.riskBlim = list(~apply(iterMeans((SB/Blim) < 1),1,max), 
    name="P3(B<Blim)", desc="Probability that SSB < Blim"),
  f.P80BMSY = list(~apply(iterMeans((SB/(Bmsy * 0.8)) > 1), 1, max),
    name="B>80Bmsy", desc="Probability that SSB > 80% x Bmsy")

)
```  


## Long-term last 25 years (75 years)



Replace deterministic MSY with MSE MSY from a run with the true, fixed $F_{MSY}$ in the  reference points

```{r}
refpts.perf = refpts
refpts.perf = rbind(refpts.perf,
                    FLPar(MSY= median(unitSums(catch(stks$Fmsy.OM[,ac((fy-10):fy)])))))

```


Compute performance metrics

```{r,warning=F,message=F}
perf <- performance(stks,
                    refpts=refpts.perf,
                    metrics=metrics, 
                    statistics=stats, 
                    years=list((fy-10):fy))

```

\pagebreak

## PLOT performance


```{r, fig.height=9,fig.width=8,warning=FALSE,message=FALSE, fig.cap = paste0("Performance: Long-term performance evaluation")}

ncol = length(unique(perf$mp)) 
pbp = plotBPs(perf,
statistics=c("a.medianFmsy","b.medianBmsy","c.medianCmsy", "d.aavC", "e.riskBlim", "f.P80BMSY"), 
size=3, target = c(a.medianFmsy=1,b.medianBmsy=1, c.medianCmsy=1,g.P80BMSY=0.8),
limit= c(e.riskBlim=0.05,c.medianCmsy=0.95),
yminmax = c(0.05, 0.95))+theme_bw()+
facet_wrap(~name,scales = "free_y",ncol=2)+
ggtitle(paste0("Performance: Reference Points"))+
ylab("Performance statistics")+
scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
theme(axis.text.x=element_blank())+xlab("Candidates")+ guides(fill= guide_legend(ncol = 1))
pbp

```


\pagebreak

### Performance Table



```{r,echo=F,message=F,warning=F}
dir.create("perftabs",showWarnings = F)

pf = perf[perf$statistic%in%statistic,]
df.pf = aggregate(data~mp+name,pf,mean)
sd.pf = aggregate(data~mp+name,pf,sd)
sd.pf = sd.pf[sd.pf$name=="Catch/MSY",] 
tab.pf = cbind(reshape2::dcast(df.pf,mp~name),SD.Catch=sd.pf$data)
tab.pf[,-1] = round(tab.pf[,-1],3)
# Add Fadv and Btrigger
tab.pf = data.frame(tab.pf,Fadv=df$Ftgt,Btrigger=df$Btrigger)

tab.pf = tab.pf[,c(1,9,7,5,8,2,6,3,4)]
write.csv(tab.pf,file="perftabs/ftune.tab.csv",row.names = F)


knitr::kable(tab.pf,"pipe",
      align ="lccccc",
      caption="Robustness tests of $F_{adv}$ in line with ICES precautionary approach under closed loop simulations with feedback control of assessment advice emulation")


```



\pagebreak

### MSE kobe plot

```{r fig5, fig.height=8,fig.width=10, fig.cap = paste0("MSE kobe plot Advice rules"), warning=FALSE}

kbcex =function(){theme(plot.title = element_text(size=10),
                        legend.key.size = unit(0.3, 'cm'), #change legend key size
                        legend.key.height = unit(0.4, 'cm'), #change legend key height
                        legend.key.width = unit(0.4, 'cm'), #change legend key width
                        legend.text = element_text(size=10)) #change legend text font size
}
kobeMPs(perf,y="a.medianFmsy", x="b.medianBmsy")+
  ylab(expression(F/F[MSY]))+xlab(expression(B/B[MSY]))+ylim(0,2.5)+kbcex()+theme()+guides(fill= guide_legend(ncol = 1))

```

